On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme

### 2020-09-11 LME-MPM.tex

\setlength{\parskip}{1cm plus 5mm minus 4mm}
%\hline
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\large
\textbf{Highlights}
\normalsize
\begin{itemize}
\item \sout{We have proposed a} A new \textcolor{green}{accurate} and efficient time integration scheme for the \acrshort{mpm} is proposed. \sout{in order to dissipate spurious stress oscillations without damping the solution.}
\item The \sout{proposal of the} \acrfull{lme} approximation technique  \sout{under the} is a robust substitute of the traditional shape functions \textcolor{green}{within the} \acrshort{mpm} framework .
\item \sout{Our} \textcolor{green}{The proposed approach might improve} \sout{have clear benefits on}  large strain fast dynamic \acrshort{mpm} simulations \sout{, where the grid-crossing pathology appears.} as grid-crossing pathology are clearly mitigated.  
\end{itemize}

%\begin{highlights}
%\end{highlights}

@@

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form
\sout{formulation} of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and \sout{an} appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Finally,  the absence of \acrshort{ugimp} values for a mesh size of 0.3325 and 0.5 millimeters in figure \eqref{fig:Dyka-error-evol} is remarkable. The reason is due to an unstable increasing \sout{of the} error suffered during \acrshort{ugimp} simulations for these mesh sizes. A feasible explanation for this phenomena could be the presence of numerical cancellation which could produce gaps between voxels. Further research should be done in this direction for getting a better comprehension of this phenomenon. Conversely, this shortcoming is not suffered by \acrshort{lme}, independently of regular or irregular nodal layout.}

@@

\textcolor{blue}{A \sout{first} comparison of the performance of both time integration schemes is presented in figure \ref{fig:Dyka-NPC-FE}. Both integration schemes show numerical oscillations with respect to the analytical solution all over the simulation time, being worse just after jump discontinuities. However, the \acrshort{npc} approach damps out these oscillations faster than the \acrshort{fe} scheme. The oscillations of the \acrshort{npc} almost disappeared before reaching subsequent jump discontinuities, while those of \acrshort{fe} do not. Moreover, oscillations of the \acrshort{fe} seems to increase further after each jump discontinuity and an unstable tendency can by foreseen. This unstable tendency is not appreciated in the \acrshort{npc} integration scheme.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparison of \acrshort{npc} and \acrshort{fe}
      performances: In the picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textcolor{blue}{Finally, an unstructured background mesh is proposed in order to validate the performance of the \acrshort{lme} shape function against the \acrshort{ugimp} one.}
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=0.8]{Figures/Velocity-structured-unstructured-lme-ugimp.png}
  }
  \caption{Sensitivity analysis of \acrshort{ugimp} and \acrshort{lme} schemes to the randomness in the layout of the background set of nodes. On the left column, sketches of the two employed meshes are depicted. On the right side, spatial distribution of the velocity magnitudes are shown for each case at second 2.66. All the simulation where carried out with the proposed \acrshort{npc} scheme.}
  \label{fig:Magnitude_velocity_impact_square_structured_unestructured}
\end{figure}
%%%%%%%%%%%%%%%%
\textcolor{blue}{ As it is observed in figure~\ref{fig:Magnitude_velocity_impact_square_structured_unestructured}, the \acrshort{lme} approximation preserves the solution even though the mesh is not structured. Meanwhile, the \acrshort{ugimp} is not able to deal with the unstructured mesh as expected. %It goes without saying that there is no point in using unstructured meshes in this type of situation. 
%Despite this, 
Thus, the \acrshort{lme} approach adopted in the present work shows a remarkable robustness \sout{in} dealing with unstructured meshes. From this last comparison it can be deduced that \acrshort{lme}  might be an adequate choice approximation technique when dealing with complex domains where a structured mesh is usually not possible .}

>>>>

\setlength{\parskip}{1cm plus 5mm minus 4mm}
%\hline
\setlength{\parskip}{1cm plus 5mm minus 4mm}
\large
\textbf{Highlights}
\normalsize
{\color{blue}
\begin{itemize}
\item A new accurate and efficient time integration scheme for the \acrshort{mpm} is proposed.
\item The \acrfull{lme} approximation technique is a robust substitute of the traditional shape functions within the \acrshort{mpm} framework .
\item The proposed approach might improve large strain fast dynamic \acrshort{mpm} simulations as grid-crossing pathology are clearly mitigated.  
\end{itemize}
}
%\begin{highlights}
%\end{highlights}

@@

In order to obtain the variational statement of the problem, let us define a
virtual displacement field such that
\begin{equation}
  \label{eq:Hilbert-space}
  \vec{u}^{\psi} \in \mathcal{H}^1_0(\Omega) = \{ \vec{u}^{\psi} \in
  \mathcal{H}^1 \mid \vec{u}^{\psi} = \vec{0}\ \text{on}\ \Gamma_d \};
\end{equation}
\textcolor{blue}{where $\vec{u}^{\psi} \in \mathcal{H}^1$ means that}
\begin{equation}
  \label{eq:cauchy-secuence}
  \Integral{3}{\vec{u}^{\psi}} < \infty\ \quad\text{and}\quad
  \Integral{3}{\tens{\varepsilon}^{\psi}} < \infty.
\end{equation}
The principle of virtual work states that the equilibrium solution to
the boundary-value problem of elasticity is the function $\vec{u} \in
\mathcal{H}^1_0$ such that, for any $\vec{u}^{\psi} \in
\mathcal{H}^1_0$,
the following holds:
%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \label{eq:BalanceMomentum_wf}
  \Integral{3}{\rho\ \left( \frac{d\vec{v}}{dt}\ - \vec{b} \right) \cdot \vec{u}^{\psi}} =
  \Integral{2}{\vec{t}\ \cdot \vec{u}^{\psi}} - \Integral{3}{\tens{\sigma} \colon
   \tens{\varepsilon}^{\psi}}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%
Thus, equation~\eqref{eq:BalanceMomentum_wf}, together with
\eqref{eq:Constitutive-equation} and
\eqref{eq:Rho-material-derivative}, represents the weak form of the problem. Next, we will discretise the set of equations of the mathematical model using a double discretisation procedure, see figure~\ref{fig:MPM-discretization}.\\

\textcolor{blue}{First, the velocity field and the virtual displacements fields are discretised in a finite set of nodes $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. An approximation of the continuum field can be obtained with the help of nodal values and appropriate interpolation functions $N_I(\vec{x})$. Also spatial derivatives of those quantities, such as gradients and divergences, are computed through the support of the background set of nodes as}

@@

\textcolor{blue}{Finally,  the absence of \acrshort{ugimp} values for a mesh size of 0.3325 and 0.5 millimeters in figure \eqref{fig:Dyka-error-evol} is remarkable. The reason is due to an unstable increasing error suffered during \acrshort{ugimp} simulations for these mesh sizes. A feasible explanation for this phenomena could be the presence of numerical cancellation which could produce gaps between voxels. Further research should be done in this direction for getting a better comprehension of this phenomenon. Conversely, this shortcoming is not suffered by \acrshort{lme}, independently of regular or irregular nodal layout.}

@@

\textcolor{blue}{A comparison of the performance of both time integration schemes is presented in figure \ref{fig:Dyka-NPC-FE}. Both integration schemes show numerical oscillations with respect to the analytical solution all over the simulation time, being worse just after jump discontinuities. However, the \acrshort{npc} approach damps out these oscillations faster than the \acrshort{fe} scheme. The oscillations of the \acrshort{npc} almost disappeared before reaching subsequent jump discontinuities, while those of \acrshort{fe} do not. Moreover, oscillations of the \acrshort{fe} seems to increase further after each jump discontinuity and an unstable tendency can by foreseen. This unstable tendency is not appreciated in the \acrshort{npc} integration scheme.} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{0.9\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Velocity_FE_vs_PCE_CFL_01}
  }
  \caption{Comparison of \acrshort{npc} and \acrshort{fe}
      performances: In the picture the velocity evolution at the point in the bar left side
    is plotted.}
  \label{fig:Dyka-NPC-FE}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\textcolor{blue}{Finally, an unstructured background mesh is proposed in order to validate the performance of the \acrshort{lme} shape function against the \acrshort{ugimp} one.}
%%%%%%%%%%%%%%%%
\begin{figure}
%\sidecaption
  \centering
  \resizebox{\hsize}{!}{
\includegraphics[scale=0.8]{Figures/Velocity-structured-unstructured-lme-ugimp.png}
  }
  \caption{Sensitivity analysis of \acrshort{ugimp} and \acrshort{lme} schemes to the randomness in the layout of the background set of nodes. On the left column, sketches of the two employed meshes are depicted. On the right side, spatial distribution of the velocity magnitudes are shown for each case at second 2.66. All the simulation where carried out with the proposed \acrshort{npc} scheme.}
  \label{fig:Magnitude_velocity_impact_square_structured_unestructured}
\end{figure}
%%%%%%%%%%%%%%%%
\textcolor{blue}{ As it is observed in figure~\ref{fig:Magnitude_velocity_impact_square_structured_unestructured}, the \acrshort{lme} approximation preserves the solution even though the mesh is not structured. Meanwhile, the \acrshort{ugimp} is not able to deal with the unstructured mesh as expected. %It goes without saying that there is no point in using unstructured meshes in this type of situation. 
%Despite this, 
Thus, the \acrshort{lme} approach adopted in the present work shows a remarkable robustness dealing with unstructured meshes. From this last comparison it can be deduced that \acrshort{lme}  might be an adequate choice approximation technique when dealing with complex domains where a structured mesh is usually not possible .}
